---
title: "analyze_ACC_discordance.Rmd"
author: "Kristen Dang"
date: "April 23, 2015"
output: html_document
---

```{r message=FALSE, echo=FALSE}
library(synapseClient)
synapseLogin()
```

## RNA-DNA discordance
```{r}
discordFile = synGet('syn3775514',downloadFile = TRUE)
discord_total = read.delim(getFileLocation(discordFile), header = TRUE)
discord_matrix = data.matrix(discord_total)
hist(discord_matrix, col = "honeydew", main = "raw discordance")

```


```{r}
sitesFile = synGet('syn3775524',downloadFile = TRUE)
sites = read.delim(getFileLocation(sitesFile), header = TRUE)
sites_matrix = data.matrix(sites)
dim(sites_matrix)
hist(sites_matrix, col = "honeydew", main = "number of sites used to calculate")
hist(sites_matrix, col = "honeydew", main = "number of sites used to calculate (zoom)", ylim = range(0,1000),breaks = 40)
```


### Making fractional discordance dataset
```{r}
metaFile = synGet("syn2299154")
metadata = read.csv(getFileLocation(metaFile), fileEncoding="macintosh")

genoIDs = metadata$DNA_genotyping_report..Genotyping.Sample_ID[match(colnames(discord_matrix), metadata$ACC_RNA_isolation..Sample.RNA.ID)]


filtered_discord = discord_matrix[match(toupper(genoIDs), toupper(rownames(discord_matrix))),]
dim(filtered_discord)
filtered_sites = sites_matrix[match(toupper(genoIDs), toupper(rownames(sites_matrix))),]
dim(filtered_sites)

fractional_discord = filtered_discord / filtered_sites
hist(fractional_discord, col = "honeydew", main = "fractional discordance")
hist(fractional_discord, col = "honeydew", xlim = range(0,0.5), breaks = 200, ylim = range(0,200), main = "discordance -- lower tail ")
```


### Self matches
```{r}
selfMatches = diag(fractional_discord)
names(selfMatches) = colnames(fractional_discord)
# How many are NA?
length(which(is.na(selfMatches)))
which(is.na(selfMatches))
# what are they supposed to match?
rownames(fractional_discord)[which(is.na(selfMatches))]

# Why are they NA?
# Which ones don't have a match between discord results and ACC RNA id? n = 4. These are truly absent from metadata file. Why? ***
colnames(discord_matrix)[which(is.na(match(colnames(discord_matrix), metadata$ACC_RNA_isolation..Sample.RNA.ID)))]
# Two others are excluded from genotyping because of missingness rate.
metadata$DNA_report..Exclude.Reason[match(colnames(discord_matrix), metadata$ACC_RNA_isolation..Sample.RNA.ID)[which(is.na(genoIDs))]]
# The rest (n = 5) match data in the metadata file, but don't have genotyping for some other reason ("FAKE" ??)
setdiff(genoIDs, rownames(fractional_discord))


hist(selfMatches, col = "honeydew", xlab = "self discordance")
hist(selfMatches, col = "honeydew", xlim = range(0,0.5), breaks = 40, ylim = range(0,150), main = "selfMatches -- low range")
hist(selfMatches, col = "honeydew", xlim = range(0.5,1), breaks = 40, ylim = range(0,20), main = "selfMatches -- high range")
# Investigating ones that don't match self
# Which RNA samples don't match self genotype
which(selfMatches > 0.5)
# Which self genotype ID are they supposed to match?
rownames(fractional_discord)[which(selfMatches > 00.5)]

# Samples with a high self-discord, but not as high as non-self values
intersect(names(which(selfMatches > 0.1)), names(which(selfMatches < 0.5)))
```



### Non-self matches
```{r}
nonSelfMatches = fractional_discord - diag(x=selfMatches)
hist(nonSelfMatches, col = "honeydew", xlab = "fractional discordance")
hist(nonSelfMatches, col = "honeydew", xlim = range(0,0.5), breaks = 200, ylim = range(0,100), main = "nonself discord -- low range")
hist(nonSelfMatches, col = "honeydew", xlim = range(0.03,0.5), breaks = 200, ylim = range(0,20))
num_low_discord = apply(nonSelfMatches,MARGIN = 2,function(x){length(intersect(which(x < 0.4), which(x > 0)))})
which(num_low_discord == 1)
suspectCols = which(num_low_discord == 1)

for (i in 1:length(suspectCols)){
  print(colnames(nonSelfMatches)[suspectCols[i]])
  print(paste('non-self matches:', intersect(names(which(nonSelfMatches[,suspectCols[i]] < 0.4)), names(which(nonSelfMatches[,suspectCols[i]] > 0))), sep = ' '))  
}

```


## RNA-RNA discordance
```{r}
allDiscordFile = synGet('syn3775542',downloadFile = TRUE)
all_discord_vec = read.delim(getFileLocation(allDiscordFile), skip = 5,header = TRUE)
all_discord_vec$fractional = as.numeric(all_discord_vec$X.2.Discordance/all_discord_vec$X.3.Number.of.sites)

all_discord_matrix = matrix(NA,nrow = 537, ncol = 537)

# reshape data into matrix
j = 1
for (i in 1:(nrow(all_discord_matrix)-1)){
  all_discord_matrix[i+1,] = c(all_discord_vec$fractional[j:(j+i-1)],rep(0,(537-i)))
  j = j+i
}

rownames(all_discord_matrix) = c("MSSM_RNA_ACC_1", as.character(unique(all_discord_vec$X.5.Sample.i)))
colnames(all_discord_matrix) = c(as.character(unique(all_discord_vec$X.6.Sample.j)), "PITT_RNA_ACC_BP_982")


hist(all_discord_matrix[lower.tri(all_discord_matrix)], col = "honeydew", main = "RNA-RNA discordance")
hist(all_discord_matrix[lower.tri(all_discord_matrix)], col = "honeydew", main = "RNA-RNA non-self discordance (zoom)", ylim = range(0,300),breaks = 40, xlab = "fractional discordance")

# Which samples are potential duplicates (e.g. high concordance)?
concordant_count = apply(all_discord_matrix,MARGIN = 2,function(x) {length(intersect(which(x < 0.2), which(x > 0)))})
names(which(concordant_count > 0))
suspectCols = which(concordant_count > 0)

for (i in 1:length(suspectCols)){
  print(colnames(all_discord_matrix)[suspectCols[i]])
  print(paste('non-self matches:', intersect(names(which(all_discord_matrix[,suspectCols[i]] < 0.2)), names(which(all_discord_matrix[,suspectCols[i]] > 0))), sep = ' '))  
}
