Functional profiles of control samples
========================================================

Load synapse and samples metadata file.
```{r cache=TRUE, warning=FALSE,results='hide'}
library('synapseClient')
synapseLogin()
dataFile = synGet('syn2340130')
metaFile = synGet('syn2299154')
```

```{r cache=TRUE}
geneCounts  = read.delim(dataFile@filePath, header=T, row.names = 1)
metadata = read.csv(getFileLocation(metaFile))
```

```{r warning=FALSE}
setwd('~/Computing/commonmind/analysis/')
source('~/Computing/sage-data-analysis/rnaseq_analysis_functions.R')
```

Make final dataset, with bad samples excluded. 

```{r}
#colnames(metadata)
metadata_freeze = metadata[-which(metadata$DLPFC_RNA_report..Exclude. == 1),]
total_samples = nrow(metadata_freeze)
table(metadata_freeze$DLPFC_RNA_report..Exclude.)
metadata_freeze$DLPFC_RNA_isolation..Sample.RNA.ID = factor(metadata_freeze$DLPFC_RNA_isolation..Sample.RNA.ID)

metadata_freeze = metadata_freeze[order(metadata_freeze$DLPFC_RNA_isolation..Sample.RNA.ID),]
geneCounts_freeze = geneCounts[,which(as.factor(colnames(geneCounts)) %in% metadata_freeze$DLPFC_RNA_isolation..Sample.RNA.ID)]
geneCounts_freeze = geneCounts_freeze[,order(as.factor(colnames(geneCounts_freeze)))]

head(colnames(geneCounts_freeze))
head(metadata_freeze$DLPFC_RNA_isolation..Sample.RNA.ID)
tail(colnames(geneCounts_freeze))
tail(metadata_freeze$DLPFC_RNA_isolation..Sample.RNA.ID)
```


Functional profiles of genes detected in control samples
--------------------------------------------------------

Make dataset: select control samples, assign Entrez IDs and exclude those without Entrez
```{r echo=FALSE, warning=FALSE}
require(goProfiles)
```
```{r}
# get control samples only
table(metadata_freeze$Dx)
controlCounts_freeze = geneCounts_freeze[,which(metadata_freeze$Dx == "Control")]

# assign Entrez ID to genes
temp = addEntrez(inCounts=controlCounts_freeze)
toRemove = which(is.na(temp$entrez))
controlCounts_entrez = temp[-toRemove,]
rm(temp, controlCounts_freeze)
```

Make three versions of the dataset where minimum reads required for detection is varied from 2:4.
```{r fig.height=4, fig.width=4}
detected_control = apply(controlCounts_entrez, MARGIN=1, FUN=countDetected)
# This plot concerns me a bit...many genes detected by very few samples only, then a large number mostly detected by all. The rest could be noise?
hist(detected_control, col = "papayawhip")

detected3_control = apply(controlCounts_entrez, MARGIN=1, FUN=countDetected, filter=3)
hist(detected3_control, col = "papayawhip")

detected4_control = apply(controlCounts_entrez, MARGIN=1, FUN=countDetected, filter=4)
hist(detected4_control, col = "papayawhip")
```

Function profile plots using GO
---------------------------------------

GO "Cellular Component" ontology, level 2
```{r,cache=TRUE}
brain.cc.profiles = basicProfile(genelist=detected_control,onto="CC",level=2,orgPackage="org.Hs.eg.db", empty.cats=F)
plotProfiles(brain.cc.profiles, "Functional profiles control samps",percentage = F, multiplePlots = F)
```

GO "Cellular Component" ontology, level 3
```{r, fig.height=8, fig.width=8,cache=TRUE}
brain.cc3.profiles = basicProfile(genelist=detected_control,onto="CC",level=3,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
#print(brain.cc3.profiles, percentage=TRUE)
brain.cc3.profiles.d4 = basicProfile(genelist=detected4_control,onto="CC",level=3,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
brain.cc3.d2d4 = mergeProfilesLists(brain.cc3.profiles, brain.cc3.profiles.d4, profNames = c("min.2", "min.4"))
plotProfiles(brain.cc3.d2d4, "Functional profiles control samps",percentage = F, multiplePlots = F, legend=TRUE)
```

GO "Cellular Component" ontology, level 4
```{r, fig.height=9, fig.width=9,cache=TRUE}
brain.cc4.profiles = basicProfile(genelist=detected_control,onto="CC",level=4,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
#save(brain.cc4.profiles,file="brain_cc4_profiles.bz2",compress="bzip2")
#load("brain_cc4_profiles.bz2")
#print(brain.cc4.profiles, percentage=TRUE)
#hist(brain.cc4.profiles[[1]]$Frequency, breaks = 20)
z = which(brain.cc4.profiles[[1]]$Frequency < 100)
brain.cc4.profiles.reduced = brain.cc4.profiles
brain.cc4.profiles.reduced[[1]] = brain.cc4.profiles.reduced[[1]][-z,]
rm(z)
plotProfiles(brain.cc4.profiles.reduced, "Major functional profiles\ncontrol samples",percentage = F, multiplePlots = F,labelWidth=30)
```

GO "Cellular Component" ontology, level 5
```{r, fig.height=9, fig.width=9,cache=TRUE}
brain.cc5.profiles = basicProfile(genelist=detected_control,onto="CC",level=5,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
#save(brain.cc5.profiles,file="brain_cc5_profiles.bz2",compress="bzip2")
#load("brain_cc5_profiles.bz2")
#print(brain.cc5.profiles, percentage=TRUE)
hist(brain.cc5.profiles[[1]]$Frequency, breaks = 20)
z = which(brain.cc5.profiles[[1]]$Frequency < 100)
brain.cc5.profiles.reduced = brain.cc5.profiles
brain.cc5.profiles.reduced[[1]] = brain.cc5.profiles.reduced[[1]][-z,]
rm(z)
plotProfiles(brain.cc5.profiles.reduced, "Major functional profiles\ncontrol samples",percentage = F, multiplePlots = F,labelWidth=30)
```

GO "Cellular Component" ontology, level 6
```{r, fig.height=9, fig.width=9,cache=TRUE}
brain.cc6.profiles = basicProfile(genelist=detected_control,onto="CC",level=6,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
#save(brain.cc6.profiles,file="brain_cc6_profiles.bz2",compress="bzip2")
#load("brain_cc6_profiles.bz2")
#print(brain.cc6.profiles, percentage=TRUE)
hist(brain.cc6.profiles[[1]]$Frequency, breaks = 20)
z = which(brain.cc6.profiles[[1]]$Frequency < 150)
brain.cc6.profiles.reduced = brain.cc6.profiles
brain.cc6.profiles.reduced[[1]] = brain.cc6.profiles.reduced[[1]][-z,]
rm(z)
plotProfiles(brain.cc6.profiles.reduced, "Major functional profiles\ncontrol samples",percentage = F, multiplePlots = F,labelWidth=30)
```


GO "Biological Process" ontology, level 2
```{r,cache=TRUE}
brain.bp.profiles = basicProfile(genelist=detected_control,onto="BP",level=2,orgPackage="org.Hs.eg.db",empty.cats=F)
plotProfiles(brain.bp.profiles, "Functional profiles\ncontrol samples",percentage = F, multiplePlots = F)
```

GO "Biological Process" ontology, level 3
```{r fig.height=12, fig.width=7,cache=TRUE}
brain.bp3.profiles = basicProfile(genelist=detected_control,onto="BP",level=3,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
#save(brain.bp3.profiles,file="brain_bp3_profiles.bz2",compress="bzip2")
#load("brain_bp3_profiles.bz2")
#print(brain.bp3.profiles, percentage=TRUE)
z = which(brain.bp3.profiles[[1]]$Frequency < 500)
brain.bp3.profiles.reduced = brain.bp3.profiles
brain.bp3.profiles.reduced[[1]] = brain.bp3.profiles.reduced[[1]][-z,]
rm(z)
#print(brain.bp3.profiles.reduced)
plotProfiles(brain.bp3.profiles.reduced, "Major functional profiles\ncontrol samples",percentage = F, multiplePlots = F, labelWidth=30)
```

GO "Biological Process" ontology, level 4
```{r,cache=TRUE}
brain.bp4.profiles = basicProfile(genelist=detected_control,onto="BP",level=4,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
#save(brain.bp4.profiles,file="brain_bp4_profiles.bz2",compress="bzip2")
#load("brain_bp4_profiles.bz2")
#print(brain.bp4.profiles, percentage=TRUE)
#hist(brain.bp4.profiles[[1]]$Frequency, breaks = 20)
z = which(brain.bp4.profiles[[1]]$Frequency < 700)
brain.bp4.profiles.reduced = brain.bp4.profiles
brain.bp4.profiles.reduced[[1]] = brain.bp4.profiles.reduced[[1]][-z,]
rm(z)
plotProfiles(brain.bp4.profiles.reduced, "Major functional profiles\ncontrol samples",percentage = F, multiplePlots = F,labelWidth=30)
```

GO "Biological Process" ontology, level 5
```{r,cache=TRUE}
brain.bp5.profiles = basicProfile(genelist=detected_control,onto="BP",level=5,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
#save(brain.bp5.profiles,file="brain_bp5_profiles.bz2",compress="bzip2")
#load("brain_bp5_profiles.bz2")
#print(brain.bp5.profiles, percentage=TRUE)
hist(brain.bp5.profiles[[1]]$Frequency, breaks = 20)
z = which(brain.bp5.profiles[[1]]$Frequency < 550)
brain.bp5.profiles.reduced = brain.bp5.profiles
brain.bp5.profiles.reduced[[1]] = brain.bp5.profiles.reduced[[1]][-z,]
rm(z)
plotProfiles(brain.bp5.profiles.reduced, "Major functional profiles\ncontrol samples",percentage = F, multiplePlots = F,labelWidth=30)
```


GO "Biological Process" ontology, level 6
```{r,cache=TRUE}
brain.bp6.profiles = basicProfile(genelist=detected_control,onto="BP",level=6,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
#save(brain.bp6.profiles,file="brain_bp6_profiles.bz2",compress="bzip2")
#load("brain_bp6_profiles.bz2")
#print(brain.bp6.profiles, percentage=TRUE)
hist(brain.bp6.profiles[[1]]$Frequency, breaks = 20)
z = which(brain.bp6.profiles[[1]]$Frequency < 500)
brain.bp6.profiles.reduced = brain.bp6.profiles
brain.bp6.profiles.reduced[[1]] = brain.bp6.profiles.reduced[[1]][-z,]
rm(z)
plotProfiles(brain.bp6.profiles.reduced, "Major functional profiles\ncontrol samples",percentage = F, multiplePlots = F,labelWidth=30)
```



GO "Molecular Function" ontology, level 2
```{r,cache=TRUE}
brain.mf.profiles = basicProfile(genelist=detected_control,onto="MF",level=2,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
plotProfiles(brain.mf.profiles, "Functional profiles control samps",percentage = F, multiplePlots = F)
# brain.mf.profiles.d4 = basicProfile(genelist=detected4_control,onto="MF",level=2,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
# brain.mf.d2d4 = mergeProfilesLists(brain.mf.profiles, brain.mf.profiles.d4, profNames = c("min.2", "min.4"))
# plotProfiles(brain.mf.d2d4, "Functional profiles control samps",percentage = F, multiplePlots = F, legend=TRUE)
```

GO "Molecular Function" ontology, level 3
```{r, fig.height=8, fig.width=8,cache=TRUE}
brain.mf3.profiles = basicProfile(genelist=detected_control,onto="MF",level=3,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
#print(brain.mf3.profiles, percentage=TRUE)
plotProfiles(brain.mf3.profiles, "Functional profiles control samps",percentage = F, multiplePlots = F)
# brain.mf3.profiles.d4 = basicProfile(genelist=detected4_control,onto="MF",level=3,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
# brain.mf3.d2d4 = mergeProfilesLists(brain.mf3.profiles, brain.mf3.profiles.d4, profNames = c("min.2", "min.4"))
# plotProfiles(brain.mf3.d2d4, "Functional profiles control samps",percentage = F, multiplePlots = F, legend=TRUE)
```

GO "Molecular Function" ontology, level 4
```{r, fig.height=10, fig.width=10,cache=TRUE}
brain.mf4.profiles = basicProfile(genelist=detected_control,onto="MF",level=4,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
#save(brain.mf4.profiles,file="brain_mf4_profiles.bz2",compress="bzip2")
#load("brain_mf4_profiles.bz2")
#print(brain.mf4.profiles, percentage=TRUE)
#hist(brain.mf4.profiles[[1]]$Frequency, breaks = 20)
z = which(brain.mf4.profiles[[1]]$Frequency < 100)
brain.mf4.profiles.reduced = brain.mf4.profiles
brain.mf4.profiles.reduced[[1]] = brain.mf4.profiles.reduced[[1]][-z,]
rm(z)
plotProfiles(brain.mf4.profiles.reduced, "Major functional profiles\ncontrol samples",percentage = F, multiplePlots = F,labelWidth=30)
```

GO "Molecular Function" ontology, level 5
```{r, fig.height=10, fig.width=10,cache=TRUE}
brain.mf5.profiles = basicProfile(genelist=detected_control,onto="MF",level=5,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
#save(brain.mf5.profiles,file="brain_mf5_profiles.bz2",compress="bzip2")
#load("brain_mf5_profiles.bz2")
#print(brain.mf5.profiles, percentage=TRUE)
hist(brain.mf5.profiles[[1]]$Frequency, breaks = 20)
z = which(brain.mf5.profiles[[1]]$Frequency < 100)
brain.mf5.profiles.reduced = brain.mf5.profiles
brain.mf5.profiles.reduced[[1]] = brain.mf5.profiles.reduced[[1]][-z,]
rm(z)
plotProfiles(brain.mf5.profiles.reduced, "Major functional profiles\ncontrol samples",percentage = F, multiplePlots = F,labelWidth=30)
```

GO "Molecular Function" ontology, level 6
```{r, fig.height=10, fig.width=10,cache=TRUE}
brain.mf6.profiles = basicProfile(genelist=detected_control,onto="MF",level=6,orgPackage="org.Hs.eg.db", empty.cats=FALSE)
#save(brain.mf6.profiles,file="brain_mf6_profiles.bz2",compress="bzip2")
#load("brain_mf6_profiles.bz2")
#print(brain.mf6.profiles, percentage=TRUE)
hist(brain.mf6.profiles[[1]]$Frequency, breaks = 20)
z = which(brain.mf6.profiles[[1]]$Frequency < 50)
brain.mf6.profiles.reduced = brain.mf6.profiles
brain.mf6.profiles.reduced[[1]] = brain.mf6.profiles.reduced[[1]][-z,]
rm(z)
plotProfiles(brain.mf6.profiles.reduced, "Major functional profiles\ncontrol samples",percentage = F, multiplePlots = F,labelWidth=30)
```