Audit figures
========================================================

```{r warning=FALSE,results='hide',echo=TRUE,message=FALSE}
library(synapseClient)
synapseLogin()
#library(lubridate)
library(ggplot2)
library(scales)
```

###Get folder ids
```{r}
qr = synQuery("select id, name from entity where parentId=='syn2495369'")
projects_files = qr$entity.id[which(qr$entity.name == "projects_files")]
user_activities = qr$entity.id[which(qr$entity.name == "users_activities")]
over_time = qr$entity.id[which(qr$entity.name == "report_over_time")]
```


## Projects

###Pie/Dot charts of current audit period composition
```{r fig.width=10}
qr = synQuery(paste("select id, name from entity where parentId=='", projects_files, "'", sep = ""))
projects_current_audit_entity = synGet(qr$entity.id[which(qr$entity.name == "project_counts_current_audit_period.txt")])
projects_current_audit = read.delim(getFileLocation(projects_current_audit_entity), sep = ":", head = FALSE)
pie(projects_current_audit$V2[c(2,3,5,6)],labels=substr(projects_current_audit$V1[c(2,3,5,6)],start=10,stop=60), main = "new projects \n current audit period", cex = 0.9)
```
```{r}
dotchart(projects_current_audit$V2[c(2,3,5,6)],labels=substr(projects_current_audit$V1[c(2,3,5,6)],start=10,stop=60),bg="red", main="new projects -- current audit", xlab = "number of projects")
```

### Data over time 
```{r}
qr = synQuery(paste("select id, name from entity where parentId=='", over_time, "'", sep = ""))
projects_time_entity = synGet(qr$entity.id[which(qr$entity.name == "projects_over_time.csv")])
projects_time = read.csv(getFileLocation(projects_time_entity), head = TRUE)
projects_creation = sapply(projects_time$CREATED_ON,function(x) {substr(x,1,10)} )
projects_creation_months = table(sapply(projects_time$CREATED_ON,function(x) {substr(x,1,7)} ))
```
#### New projects bar plot
```{r}
projects_creation_months_df = as.data.frame(projects_creation_months)
projects_creation_months_df$date = as.Date(paste(projects_creation_months_df[,1], "-28", sep=""))
ggplot(data=projects_creation_months_df, aes(x=date, y=Freq)) + geom_bar(stat="identity") + ggtitle("New Projects") + ylab("Number of new projects") + xlab("Date")
```
#### Total projects over time line chart
```{r}
total_projects = rep(NA,length(projects_creation_months))
names(total_projects) = names(projects_creation_months)
total_projects[1] = projects_creation_months[1]
for (i in 2:length(projects_creation_months)) {
  total_projects[i] = total_projects[i-1] + projects_creation_months[i]
}
total_projects_df = as.data.frame(total_projects)
total_projects_df$date = as.Date(paste(rownames(total_projects_df), "-28", sep=""))
qplot(x=date, y = total_projects, data=total_projects_df, geom="line", main = "Number of Projects", ylab = "Total projects", xlab = "Date")
#last_plot() + scale_x_date(breaks = date_breaks("3 months"))
```




## Files


###Pie/Dot charts of current audit period composition
```{r fig.width=10}
qr = synQuery(paste("select id, name from entity where parentId=='", projects_files, "'", sep = ""))
files_current_audit_entity = synGet(qr$entity.id[which(qr$entity.name == "file_counts_current_audit_period.txt")])
files_current_audit = read.delim(getFileLocation(files_current_audit_entity), sep = ":", head = FALSE)
pie(files_current_audit$V2[c(2,3,5,6)],labels=substr(files_current_audit$V1[c(2,3,5,6)],start=10,stop=60), main = "new files \n current audit period", cex = 0.9)

colnames(files_current_audit) = c("names", "counts")
x = ggplot(files_current_audit[c(2,3,5,6),], aes(x = factor(names),y = counts, fill = counts))

x + geom_bar()
x + geom_bar(stat = "identity")
x + coord_polar

pie = ggplot(files_current_audit[c(2,3,5,6),], aes(x = names, fill = counts)) + geom_bar(width = 1, stat="identity")
pie + coord_polar(theta = "y")
```
```{r}
dotchart(files_current_audit$V2[c(2,3,5,6)],labels=substr(files_current_audit$V1[c(2,3,5,6)],start=10,stop=60),bg="red", main="new files -- current audit period", xlab = "number of files")
```


###Line charts of total files over time 
```{r}
qr = synQuery(paste("select id, name from entity where parentId=='", over_time, "'", sep = ""))
files_time_entity = synGet(qr$entity.id[which(qr$entity.name == "public_files_over_time.csv")])
files_time = read.csv(getFileLocation(files_time_entity), head = TRUE)
files_creation = sapply(files_time$CREATED_ON,function(x) {substr(x,1,10)} )
files_creation_months = table(sapply(files_time$CREATED_ON,function(x) {substr(x,1,7)} ))
```
New public files plot
```{r}
files_creation_months_df = as.data.frame(files_creation_months)
files_creation_months_df$date = as.Date(paste(files_creation_months_df[,1], "-28", sep=""))
ggplot(data=files_creation_months_df, aes(x=date, y=Freq)) + geom_bar(stat="identity") + ggtitle("New Public Files") + ylab("Number of new public files") + xlab("Date")
```
Total public files over time
```{r}
total_files = rep(NA,length(files_creation_months))
names(total_files) = names(files_creation_months)
total_files[1] = files_creation_months[1]
for (i in 2:length(files_creation_months)) {
  total_files[i] = total_files[i-1] + files_creation_months[i]
}
total_files_df = as.data.frame(total_files)
total_files_df$date = as.Date(paste(rownames(total_files_df), "-28", sep=""))
qplot(x=date, y = total_files, data=total_files_df, geom="line", main = "Number of Public Files", ylab = "Total public files", xlab = "Date")
#last_plot() + scale_x_date(breaks = date_breaks("3 months"))
```



### Which projects have controlled files?
```{r}
qr = synQuery(paste("select id, name from entity where parentId=='", projects_files, "'", sep = ""))
public_file_controlled_entity = synGet(qr$entity.id[grep("public_files_controlled_snapshot", qr$entity.name)])
public_file_controlled = read.csv(getFileLocation(public_file_controlled_entity))
table(public_file_controlled$PROJECT_NAME)
```


### Other plots? Not sure what would be helpful to see from this data...?
```{r}
public_file_uploads_ca_entity = synGet(qr$entity.id[grep("public_file_uploads_sage_current_audit_period.csv", qr$entity.name)])
public_file_uploads_ca = read.csv(getFileLocation(public_file_uploads_ca_entity))
head(public_file_uploads_ca)
```



## Users


### Active users -- days on synapse
```{r}
qr = synQuery(paste("select id, name from entity where parentId=='", user_activities, "'", sep = ""))
nonsage_users_entity = synGet(qr$entity.id[which(qr$entity.name == "active_non_sage_users.csv")])
nonsage_users = read.csv(getFileLocation(nonsage_users_entity), head = TRUE)
head(nonsage_users)
hist(nonsage_users$DAYS, col = "powderblue", main = "active non-Sage users", xlab = "number of days using Synapse during audit period", breaks = 20)
sage_users_entity = synGet(qr$entity.id[which(qr$entity.name == "active_sage_users.csv")])
sage_users = read.csv(getFileLocation(sage_users_entity), head = TRUE)
head(sage_users)
hist(sage_users$DAYS, col = "powderblue", main = "active Sage users", xlab = "number of days using Synapse during audit period", breaks = 20)
```

Not sure what to do with this data...?
```{r}
user_details_prj_entity = synGet(qr$entity.id[which(qr$entity.name == "user_details_projects.csv")])
user_details_prj = read.csv(getFileLocation(user_details_prj_entity), head = TRUE)
head(user_details_prj)
```



###Get users over time data
```{r}
qr = synQuery(paste("select id, name from entity where parentId=='", over_time, "'", sep = ""))
users_time_entity = synGet(qr$entity.id[which(qr$entity.name == "non_sage_users_over_time.csv")])
users_time = read.csv(getFileLocation(users_time_entity), head = TRUE)
creation = sapply(users_time$CREATED_ON,function(x) {substr(x,1,10)} )
creation_months = table(sapply(users_time$CREATED_ON,function(x) {substr(x,1,7)} ))
```
#### New users bar plot
```{r}
creation_months_df = as.data.frame(creation_months)
creation_months_df$date = as.Date(paste(creation_months_df[,1], "-28", sep=""))
#ggplot(data=creation_months_df, aes(x=date, y=Freq)) + geom_bar(stat="identity") + ggtitle("New Users") + ylab("Number of new users") + xlab("Date")
creation_asdate = as.data.frame(as.Date(creation))
colnames(creation_asdate) = c("date")
ggplot(data=creation_asdate, aes(x=date)) + geom_bar(stat="bin") + ggtitle("New Users") + ylab("Number of new users") + xlab("Date") + scale_x_date(labels = date_format("%m/%d"), breaks = date_breaks("week"), limits = c(as.Date("2014-02-1"), as.Date("2014-04-30")))
```

#### Total users over time
```{r}
total_users = rep(NA,length(creation_months))
names(total_users) = names(creation_months)
total_users[1] = creation_months[1]
for (i in 2:length(creation_months)) {
  total_users[i] = total_users[i-1] + creation_months[i]
}
total_users_df = as.data.frame(total_users)
total_users_df$date = as.Date(paste(rownames(total_users_df), "-28", sep=""))
qplot(x=date, y = total_users, data=total_users_df, geom="line", main = "Number of Users", ylab = "Total users", xlab = "Date")
#last_plot() + scale_x_date(breaks = date_breaks("3 months"))
```
