---
title: "clustering_vargenes_nosex_model.Rmd"
author: "Kristen Dang"
date: "August 17, 2015"
output: html_document
---


```{r message=FALSE,warning=FALSE}
library('rGithubClient')
source('/Users/kristen/Computing/external_software/rgithubclient_authenticate.R')
```

```{r message=FALSE}
library('gplots')
library('MASS')
library(ConsensusClusterPlus)
sourceRepoFile(sageCode, 'scri_cran/cranio_common_routines.R')
setwd('~/Computing/cranio/')
cutoff = 0.05
```


## Get filtered, residualized, case-only dataset
### Using residuals of model that does NOT include AlkP or sample_type
```{r}
residEntity = synGet('syn4893932')
resid = data.matrix(read.delim(getFileLocation(residEntity),row.names=1))
boxplot(resid,outline = FALSE)
resid.cor = as.dist(cor(resid))
resid_mc = sweep(resid,1, apply(resid,1,median,na.rm=T)) # median-center
hist(cor(resid))
hist(cor(resid_mc))

colnames(resid) = sub("X",replacement = "",x = colnames(resid))
colnames(resid) = gsub("[.]",replacement = "-",x = colnames(resid))
```


## Identify variable genes, exclude sex-chromosome genes
```{r}
Ygenes = getBM(attributes=c("ensembl_gene_id"),filters="chromosome_name",values="Y", mart=Hs)
Xgenes = getBM(attributes=c("ensembl_gene_id"),filters="chromosome_name",values="X", mart=Hs)
detach("package:lubridate", unload=TRUE)


dim(resid)
fitEnt = synGet('syn4893933')
load(getFileLocation(fitEnt))
hist(fit$sigma, col = "honeydew",breaks = 20)
truehist(fit$sigma, col = "honeydew")
sum(fit$sigma > 2)
sum(fit$sigma > 1.5)
sum(fit$sigma > 3)
sum(fit$sigma > 4)
highvar2 = setdiff(names(fit$Amean)[which(fit$sigma > 2)], union(Xgenes[,1], Ygenes[,1]))
highvar1.5 = setdiff(names(fit$Amean)[which(fit$sigma > 1.5)], union(Xgenes[,1], Ygenes[,1]))
highvar3 = setdiff(names(fit$Amean)[which(fit$sigma > 3)], union(Xgenes[,1], Ygenes[,1]))
highvar4 = setdiff(names(fit$Amean)[which(fit$sigma > 4)], union(Xgenes[,1], Ygenes[,1]))


```


## consensus clustering
```{r echo=TRUE,eval=FALSE}
setwd('/Users/kristen/Computing/cranio/consensus_cluster_genes/alt_age_IGDD_model/')
results_hc = ConsensusClusterPlus(d=resid[(rownames(resid) %in% highvar2),],maxK = 35,reps = 1000,clusterAlg = "hc",distance = "spearman",pFeature = 0.9,writeTable=TRUE,plot="pdf",title="consCluster_resid_hc_high2_spear_pF9")
save(results_hc,file = "variable_genes_results_hc.Robj.bz2",compress = "bzip2")
test = results_hc[1:20]
test2 = calcICL(test,title = 'variable_genes_results_hc_k9h2',plot = 'pdf',writeTable = TRUE)

results_hc = ConsensusClusterPlus(d=resid[(rownames(resid) %in% highvar4),],maxK = 35,reps = 1000,clusterAlg = "hc",distance = "spearman",pFeature = 0.5,writeTable=TRUE,plot="pdf",title="consCluster_resid_hc_high4_spear_pF5")
save(results_hc,file = "variable_genes_results_hc_p5hv4.Robj.bz2",compress = "bzip2")
results_hc = ConsensusClusterPlus(d=resid[(rownames(resid) %in% highvar1.5),],maxK = 35,reps = 1000,clusterAlg = "hc",distance = "spearman",pFeature = 0.9,writeTable=TRUE,plot="pdf",title="consCluster_resid_hc_high15_spear_pF9")
save(results_hc,file = "variable_genes_results_hc_p9hv15.Robj.bz2",compress = "bzip2")
results_hc = ConsensusClusterPlus(d=resid[(rownames(resid) %in% highvar2),],maxK = 35,reps = 1000,clusterAlg = "hc",distance = "spearman",pFeature = 0.7,writeTable=TRUE,plot="pdf",title="consCluster_resid_hc_high2_spear_pF7")
save(results_hc,file = "variable_genes_results_hc_p7hv2.Robj.bz2",compress = "bzip2")

# varying number input genes
# pFeature = 0.7
results_pam = ConsensusClusterPlus(d=resid[(rownames(resid) %in% highvar1.5),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.7,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high15_spear_pF7") 
results_pam = ConsensusClusterPlus(d=resid[(rownames(resid) %in% highvar2),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.7,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high2_spear_pF7") 
results_pam = ConsensusClusterPlus(d=resid[(rownames(resid) %in% highvar4),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.7,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high4_spear_pF7") 


# pFeature = 0.9
results_pam = ConsensusClusterPlus(d=resid[(rownames(resid) %in% highvar1.5),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.9,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high15_spear_pF9") 
results_pam = ConsensusClusterPlus(d=resid[(rownames(resid) %in% highvar2),],maxK = 35,reps = 5000,clusterAlg = "pam",distance = "spearman",pFeature = 0.9,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high2_spear_pF9_5krep") 
save(results_pam,file = "variable_genes_results_pam_k9h2_5kreps.Robj.bz2",compress = "bzip2")
test = results_pam[1:20]
test2 = calcICL(test,title = 'variable_genes_results_pam_k9hr5k2',plot = 'pdf',writeTable = TRUE)


results_pam = ConsensusClusterPlus(d=resid[(rownames(resid) %in% highvar4),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.9,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high4_spear_pF9") 
save(results_pam,file = "variable_genes_results_pam_p9hv4.Robj.bz2",compress = "bzip2")

# pFeature = 0.5
results_pam = ConsensusClusterPlus(d=resid[(rownames(resid) %in% highvar1.5),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.5,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high15_spear_pF5") 
results_pam = ConsensusClusterPlus(d=resid[(rownames(resid) %in% highvar2),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.5,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high2_spear_pF5") 
results_pam = ConsensusClusterPlus(d=resid[(rownames(resid) %in% highvar4),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.5,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high4_spear_pF5") 
```


## Compare groups
```{r}
# Joint heatmap of hclust consensus and pclust consensus
load("/Users/kristen/Computing/cranio/consensus_cluster_genes/alt_age_IGDD_model/variable_genes_results_hc.Robj.bz2")
load('/Users/kristen/Computing/cranio/consensus_cluster_genes/alt_age_IGDD_model/variable_genes_results_pam_k9h2_5kreps.Robj.bz2')

varGenes_combined.classes = cbind(results_hc[[19]]$consensusClass,results_pam[[11]]$consensusClass)
colnames(varGenes_combined.classes) = c("hierarchy", "partition")
varGenes_ordered = varGenes_combined.classes[order(varGenes_combined.classes[,1]),]
heatmap(varGenes_ordered, col = rainbow(19),Colv = NA,scale = "none",margins = c(3,3),cexCol = 1,Rowv = NA)
varGenes_ordered = varGenes_combined.classes[order(varGenes_combined.classes[,2]),]
heatmap(varGenes_ordered, col = rainbow(19),Colv = NA,scale = "none",margins = c(3,3),cexCol = 1,Rowv = NA)
```


# Identify well-defined groups
```{r}
# From pam results
pamICI = read.csv('/Users/kristen/Computing/cranio/consensus_cluster_genes/alt_age_IGDD_model/variable_genes_results_pam_k9hr5k/variable_genes_results_pam_k9hr5k2.summary.cluster.consensus.csv')
selected = pamICI[pamICI$k == 11,]
barplot(selected$clusterConsensus,names.arg = seq(1,11))
abline(0.75,0,col = "red", lty = 2)
selectedGroups = which(selected$clusterConsensus > 0.74)
pamSamplesInSelectedGroups = names(results_pam[[11]]$consensusClass[results_pam[[11]]$consensusClass %in% selectedGroups])
pam_selected_groups = results_hc[[19]]$consensusClass[results_hc[[19]]$consensusClass %in% selectedGroups]
write.table(pam_selected_groups,file = "pam_selected_groups.tsv",quote = FALSE,sep = "\t",row.names = TRUE,col.names = FALSE)
synStore(File(path = "pam_selected_groups.tsv",parentId='syn4893931'))

# From hclust results
hcICI = read.csv('/Users/kristen/Computing/cranio/consensus_cluster_genes/alt_age_IGDD_model/variable_genes_results_hc_k9h2/variable_genes_results_hc_k9h2.summary.cluster.consensus.csv')
selected = hcICI[hcICI$k == 19,]
barplot(selected$clusterConsensus,names.arg = seq(1,19))
abline(0.75,0,col = "red", lty = 2)
selectedGroups = which(selected$clusterConsensus > 0.74)
hcSamplesInSelectedGroups = names(results_hc[[19]]$consensusClass[results_hc[[19]]$consensusClass %in% selectedGroups])
hc_selected_groups = results_hc[[19]]$consensusClass[results_hc[[19]]$consensusClass %in% selectedGroups]
write.table(hc_selected_groups,file = "hc_selected_groups.tsv",quote = FALSE,sep = "\t",row.names = TRUE,col.names = FALSE)
synStore(File(path = "hc_selected_groups.tsv",parentId='syn4893931'))

```


Eliminate genes with very low expression, try clustering again. 
(Results show little difference in clustering.)
```{r eval=FALSE}
dataEnt = synGet('syn4738853')
load(getFileLocation(dataEnt))

hist(cpm(in.dge,normalized.lib.sizes = TRUE,log = TRUE))
depthNorm = cpm(in.dge,normalized.lib.sizes = TRUE,log = TRUE)
lowGenes = apply(depthNorm,MARGIN = 1,function(z){length(which(z < 0))})
hist(lowGenes/202,breaks = 400,xlim = c(0.8,1),ylim = c(0,2000))
length(which(lowGenes/202 > 0.97 ))
filtered.dge = in.dge[-which(lowGenes/202 > 0.97 ),]

sourceRepoFile(sageCode, "scri_cran/process_metadata.R")
metadataFiltered = processMetadata(ver = 5,plot = FALSE)
metadataFiltered$Sample_Type = as.character(metadataFiltered$Sample_Type)
metadataFiltered$Sample_Type[grep("Coronal",metadataFiltered$Sample_Type)] = "Coronal"
metadataFiltered$Sample_Type[grep("Metopic",metadataFiltered$Sample_Type)] = "Metopic"
metadataFiltered$Sample_Type[grep("Sagittal",metadataFiltered$Sample_Type)] = "Sagittal"
metadataFiltered$Sample_Type = as.factor(metadataFiltered$Sample_Type)
minimalSetNoSex = c("Age_mos.","PCT_CORRECT_STRAND_READS","PCT_INTRONIC_BASES", "Initial_growth_duration_days")

caseMeta = metadataFiltered[-which(metadataFiltered$Sample_Type == "Control"),]
dim(caseMeta)
tempMeta = caseMeta[caseMeta$Px_Code %in% colnames(filtered.dge),]
dim(tempMeta)

## Run model and get residuals
tempMeta$Sample_Type = as.factor(as.character(tempMeta$Sample_Type))
tempModel = model.matrix(as.formula(paste("~",paste(minimalSetNoSex,collapse = "+"),sep = "")), data = tempMeta)

data.voom = voom(filtered.dge,tempModel,plot=FALSE) 
fit = lmFit(data.voom,tempModel)
save(fit,file = "resid_cases_pIGDD_nolow_fit.Robj.bz2",compress = "bzip2")
resid_nolow = residuals(fit,y = data.voom)
write.table(formatC(resid,digits=6,format="fg"),file = "scri-cran_resid_cases_pIGDD_nolow.tsv",quote = FALSE,sep = "\t",row.names = TRUE,col.names = TRUE)

## Save data to Synapse
synStore(File(path = "scri-cran_resid_cases_pIGDD_nolow.tsv",parentId='syn4893931'))
synStore(File(path = "resid_cases_pIGDD_nolow_fit.Robj.bz2",parentId='syn4893931'))


dim(resid_nolow)
hist(resid_nolow)
hist(fit$sigma, col = "honeydew",breaks = 20)
truehist(fit$sigma, col = "honeydew")
sum(fit$sigma > 2)
sum(fit$sigma > 1.5)
sum(fit$sigma > 3)
sum(fit$sigma > 4)
highvar2 = setdiff(names(fit$Amean)[which(fit$sigma > 2)], union(Xgenes[,1], Ygenes[,1]))
highvar1.5 = setdiff(names(fit$Amean)[which(fit$sigma > 1.5)], union(Xgenes[,1], Ygenes[,1]))
highvar3 = setdiff(names(fit$Amean)[which(fit$sigma > 3)], union(Xgenes[,1], Ygenes[,1]))
highvar4 = setdiff(names(fit$Amean)[which(fit$sigma > 4)], union(Xgenes[,1], Ygenes[,1]))

results_hc = ConsensusClusterPlus(d=resid_nolow[(rownames(resid_nolow) %in% highvar2),],maxK = 35,reps = 1000,clusterAlg = "hc",distance = "spearman",pFeature = 0.9,writeTable=TRUE,plot="pdf",title="consCluster_resid_hc_h2_spear_pF9_nolow") # no qualitative difference from including all genes.
save(results_hc,file = "variable_genes_results_hch2p9_nolow.Robj.bz2",compress = "bzip2")
test = results_hc[1:25]
test2 = calcICL(test,title = 'variable_genes_results_hck9h2_nolow',plot = 'pdf',writeTable = TRUE)


results_pam = ConsensusClusterPlus(d=resid_nolow[(rownames(resid_nolow) %in% highvar2),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.9,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_h2_spear_pF9_nolow") 
save(results_pam,file = "variable_genes_results_pamk9h2_nolow.Robj.bz2",compress = "bzip2")
test = results_pam[1:20]
test2 = calcICL(test,title = 'variable_genes_results_pam_k9hr_nolow',plot = 'pdf',writeTable = TRUE)
