---
title: "consensus_clustering_variable_genes.Rmd"
author: "Kristen Dang"
date: "June 18, 2015"
output: html_document
---


```{r message=FALSE,warning=FALSE}
library('rGithubClient')
library('gplots')
library('MASS')
library(ConsensusClusterPlus)


token = read.delim('~/Movies/rGHclient_token.txt',header = FALSE)
setGithubToken(as.character(token[1,1]))
sageCode = getRepo(repository="kkdang/sage-data-analysis")
sourceRepoFile(sageCode, 'scri_cran/cranio_common_routines.R')
proteinCoding_genes = getByBiotype()
setwd('~/Computing/cranio/')

cutoff = 0.05

```



## Get filtered, residualized, case-only dataset
### Using residuals of model that does NOT include AlkP or sample_type
```{r}
residEntity = synGet('syn4586436')
resid = read.delim(getFileLocation(residEntity),row.names=1)
boxplot(resid,outline = FALSE)
resid.cor = as.dist(cor(resid))
resid_mc = sweep(resid,1, apply(resid,1,median,na.rm=T)) # makes no difference to consensus cluster using PAM
hist(cor(resid))
hist(cor(resid_mc))

colnames(resid) = sub("X",replacement = "",x = colnames(resid))
colnames(resid) = gsub("[.]",replacement = "-",x = colnames(resid))
```


## Identify variable genes
```{r}
dim(resid)
fitEnt = synGet('syn4636650')
load(getFileLocation(fitEnt))
hist(fit$sigma, col = "honeydew",breaks = 20)
truehist(fit$sigma, col = "honeydew")
sum(fit$sigma > 2)
sum(fit$sigma > 1.5)
sum(fit$sigma > 3)
sum(fit$sigma > 4)
highvar2 = names(fit$Amean)[which(fit$sigma > 2)]
highvar1.5 = names(fit$Amean)[which(fit$sigma > 1.5)]
highvar3 = names(fit$Amean)[which(fit$sigma > 3)]
highvar4 = names(fit$Amean)[which(fit$sigma > 4)]
```


## consensus clustering
```{r echo=TRUE,eval=FALSE}
setwd('~/Computing/cranio/consensus_cluster_genes/')
plot(1, main = "HC consensus")
results_hc = ConsensusClusterPlus(d=resid[which(rownames(resid) %in% highvar2),],maxK = 35,reps = 1000,clusterAlg = "hc",distance = "spearman",pFeature = 0.9,writeTable=TRUE,plot="pdf",title="consCluster_resid_hc_high2_spear_pF9")
save(results_hc,file = "variable_genes_results_hc.Robj.bz2",compress = "bzip2")

# varying number input genes
# pFeature = 0.7
results_pam = ConsensusClusterPlus(d=resid[which(rownames(resid) %in% highvar1.5),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.7,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high15_spear_pF7") # strongly supports 2 groups
results_pam = ConsensusClusterPlus(d=resid[which(rownames(resid) %in% highvar2),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.7,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high2_spear_pF7") # strongly supports 2 groups
results_pam = ConsensusClusterPlus(d=resid[which(rownames(resid) %in% highvar4),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.7,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high4_spear_pF7") # strongly supports 2 groups


# pFeature = 0.9
results_pam = ConsensusClusterPlus(d=resid[which(rownames(resid) %in% highvar1.5),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.9,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high15_spear_pF9") # strongly supports 2 groups
results_pam = ConsensusClusterPlus(d=resid[which(rownames(resid) %in% highvar2),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.9,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high2_spear_pF9") # strongly supports 2 groups
save(results_pam,file = "variable_genes_results_pam.Robj.bz2",compress = "bzip2")

results_pam = ConsensusClusterPlus(d=resid[which(rownames(resid) %in% highvar4),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.9,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high4_spear_pF9") # strongly supports 2 groups


# pFeature = 0.5
results_pam = ConsensusClusterPlus(d=resid[which(rownames(resid) %in% highvar1.5),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.5,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high15_spear_pF5") # strongly supports 2 groups
results_pam = ConsensusClusterPlus(d=resid[which(rownames(resid) %in% highvar2),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.5,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high2_spear_pF5") # strongly supports 2 groups
results_pam = ConsensusClusterPlus(d=resid[which(rownames(resid) %in% highvar4),],maxK = 35,reps = 1000,clusterAlg = "pam",distance = "spearman",pFeature = 0.5,writeTable=TRUE,plot="pdf",title="consCluster_resid_pam_high4_spear_pF5") # strongly supports 2 groups
```


## Compare groups
```{r}
# Joint heatmap of hclust consensus and pclust consensus
load("~/Computing/cranio/variable_genes_results_hc.Robj.bz2")
load('~/Computing/cranio/variable_genes_results_pam.Robj.bz2')

varGenes_combined.classes = cbind(results_hc[[2]]$consensusClass,results_pam[[2]]$consensusClass)
colnames(varGenes_combined.classes) = c("hierarchy", "partition")
heatmap(varGenes_combined.classes, col = rainbow(2),Colv = NA,scale = "none",margins = c(3,3),cexCol = 1)

varGenes_combined.classes = cbind(results_hc[[7]]$consensusClass,results_pam[[7]]$consensusClass)
colnames(varGenes_combined.classes) = c("hierarchy", "partition")
heatmap(varGenes_combined.classes, col = rainbow(7),Colv = NA,scale = "none",margins = c(3,3),cexCol = 1)

varGenes_combined.classes = cbind(results_hc[[11]]$consensusClass,results_pam[[11]]$consensusClass)
colnames(varGenes_combined.classes) = c("hierarchy", "partition")
heatmap(varGenes_combined.classes, col = rainbow(11),Colv = NA,scale = "none",margins = c(3,3),cexCol = 1)

varGenes_combined.classes = cbind(results_hc[[15]]$consensusClass,results_pam[[15]]$consensusClass)
colnames(varGenes_combined.classes) = c("hierarchy", "partition")
heatmap(varGenes_combined.classes, col = rainbow(15),Colv = NA,scale = "none",margins = c(3,3),cexCol = 1)
```




## Check whether this grouping corresponds to any metadata
```{r}
varGenes_combined.classes = cbind(results_hc[[2]]$consensusClass,results_pam[[2]]$consensusClass)
colnames(varGenes_combined.classes) = c("hierarchy", "partition")

# Make case metadata object
caseMeta = metadata[metadata$Px_Code %in% colnames(resid), ]
caseMeta$Sample_Type = as.character(caseMeta$Sample_Type)
caseMeta$Sample_Type[grep("Coronal",caseMeta$Sample_Type)] = "Coronal"
caseMeta$Sample_Type[grep("Metopic",caseMeta$Sample_Type)] = "Metopic"
caseMeta$Sample_Type[grep("Sagittal",caseMeta$Sample_Type)] = "Sagittal"
caseMeta$Sample_Type = as.factor(caseMeta$Sample_Type)



# Enriched for sample type?
table(caseMeta$Sample_Type[which(caseMeta$Px_Code %in% names(which(varGenes_combined.classes[,2] == as.factor(1))))])
table(caseMeta$Sample_Type[which(caseMeta$Px_Code %in% names(which(varGenes_combined.classes[,2] == as.factor(2))))])
x = as.vector(table(caseMeta$Sample_Type[which(caseMeta$Px_Code %in% names(which(varGenes_combined.classes[,2] == as.factor(1))))]))
y = rbind(x, as.vector(table(caseMeta$Sample_Type[which(caseMeta$Px_Code %in% names(which(varGenes_combined.classes[,2] == as.factor(2))))])))
chisq.test(as.numeric(y),simulate.p.value = TRUE)


# library Batch
table(caseMeta$libraryBatch[which(caseMeta$Px_Code %in% names(which(varGenes_combined.classes[,2] == as.factor(1))))])
table(caseMeta$libraryBatch[which(caseMeta$Px_Code %in% names(which(varGenes_combined.classes[,2] == as.factor(2))))])

# Sample Origin
table(caseMeta$Sample_Origin[which(caseMeta$Px_Code %in% names(which(varGenes_combined.classes[,2] == as.factor(1))))])
table(caseMeta$Sample_Origin[which(caseMeta$Px_Code %in% names(which(varGenes_combined.classes[,2] == as.factor(2))))])

# Sex
table(caseMeta$Sex[which(caseMeta$Px_Code %in% names(which(varGenes_combined.classes[,2] == as.factor(1))))])
table(caseMeta$Sex[which(caseMeta$Px_Code %in% names(which(varGenes_combined.classes[,2] == as.factor(2))))])


# Scan Date
table(caseMeta$ScanDate[which(caseMeta$Px_Code %in% names(which(varGenes_combined.classes[,2] == as.factor(1))))])
table(caseMeta$ScanDate[which(caseMeta$Px_Code %in% names(which(varGenes_combined.classes[,2] == as.factor(2))))])


```


## Characterize groups, for 2 groups
```{r}
# Using random forest, only on variable genes
library('randomForest')
x = resid[which(rownames(resid) %in% highvar2),]
varGenes.rf = randomForest(x=t(x),y=as.factor(varGenes_combined.classes[,2]),importance=TRUE, proximity=TRUE, ntree=1000)
imp = round(importance(varGenes.rf), 2)
barplot(abs(imp[order(abs(imp[,3]),decreasing=TRUE),3]), las = 2, main = "MeanDecAcc",names.arg = "")
barplot(imp[order(abs(imp[,3]),decreasing=TRUE),3], las = 2, main = "MeanDecAcc",names.arg = "")
barplot(imp[order(imp[,4],decreasing=TRUE),4], las = 2, main = "IncNodePurity",names.arg = "") 
varImpPlot(varGenes.rf)
varImpPlot(varGenes.rf,n.var = 150)
print(varGenes.rf)
varGenes.centers = classCenter(x=t(x),label=as.factor(varGenes_combined.classes[,2]),prox=varGenes.rf$proximity)
heatmap.2(varGenes.centers,labCol="",trace='n',col=bluered,scale='r')

head(imp)
length(which(imp[,3] > 4))
names(which(imp[,3] > 4))
write.table(rownames(imp),file = "vargenes_list.txt",quote = FALSE,sep = "\n",row.names = FALSE,col.names = FALSE)

# Standard DE of 2 groups, using all PALX20 genes
# starting with residuals of previous model
groupStatus = as.factor(varGenes_combined.classes[,2])
minimalSet = c("Sex","Age_mos.","PCT_CORRECT_STRAND_READS","PCT_INTRONIC_BASES")
tempModel = model.matrix(~groupStatus)
fit = lmFit(resid,tempModel)
fit = eBayes(fit)
DEgeneTable = topTable(fit,coef = "groupStatus2",number = nrow(resid))
xSig = rownames(DEgeneTable)[DEgeneTable$adj.P.Val < 0.01]
xStatus = as.numeric(rownames(fit) %in% xSig)
plotMA(fit,status = xStatus,cex=c(0.3,0.5),legend = FALSE,col=c("black","green3"),values=c(0,1))

# starting with full model, including groupStatus with it 
tempMeta = caseMeta
tempMeta$Sample_Type = as.factor(as.character(tempMeta$Sample_Type))
tempModel = model.matrix(as.formula(paste("~groupStatus",paste(minimalSet,collapse = "+"),sep = "+")), data = tempMeta)
dgeEnt = synGet('syn4738853')
load(getFileLocation(dgeEnt))
data.voom = voom(in.dge,tempModel,plot=FALSE) 
fit = lmFit(data.voom,tempModel)
fit = eBayes(fit)
DEgeneTable = topTable(fit,coef = "groupStatus2",number = nrow(resid))
xSig = rownames(DEgeneTable)[DEgeneTable$adj.P.Val < 0.01]
xStatus = as.numeric(rownames(fit) %in% xSig)
plotMA(fit,status = xStatus,cex=c(0.3,0.5),legend = FALSE,col=c("black","green3"),coef = "groupStatus2",values=c(0,1))


# starting with full model, including groupStatus with it, using only variable genes
tempModel = model.matrix(as.formula(paste("~groupStatus",paste(minimalSet,collapse = "+"),sep = "+")), data = tempMeta)
vargenes.dge = in.dge[which(rownames(in.dge) %in% highvar2),]
data.voom = voom(vargenes.dge,tempModel,plot=FALSE) 
fit = lmFit(data.voom,tempModel)
fit = eBayes(fit)
DEgeneTable = topTable(fit,coef = "groupStatus2",number = nrow(resid))
xSig = rownames(DEgeneTable)[DEgeneTable$adj.P.Val < 0.01]
xStatus = as.numeric(rownames(fit) %in% xSig)
plotMA(fit,status = xStatus,cex=c(0.3,0.5),legend = FALSE,col=c("black", "green3"),coef = "groupStatus2",values=c(0,1))
```


# simulation -- are there other 2-group assignments with as many significant genes?
```{r echo=TRUE, eval=FALSE}
simSig = rep(NA,1000)
for (i in 1:1000) {
  simStatus = as.factor(rbinom(n = 202,size = 1,prob = 0.5)+1)
  tempModel = model.matrix(as.formula(paste("~simStatus",paste(minimalSet,collapse = "+"),sep = "+")), data = tempMeta)
  data.voom = voom(vargenes.dge,tempModel,plot=FALSE) 
  fit = lmFit(data.voom,tempModel)
  fit = eBayes(fit)
  DEgeneTable = topTable(fit,coef = "simStatus2",number = nrow(vargenes.dge))
  simSig[i] = length(which(DEgeneTable$adj.P.Val < 0.01))
}
pdf(file = "hist_sig-genes_random2grp.pdf")
hist(simSig, breaks = 20, col = "lawngreen", main = "number of sig genes\nfor random 2 group assignment", xlab = "number of significant genes")
dev.off()

```



## Characterize for 10 groups
```{r}
varGenes_combined.classes = cbind(results_hc[[10]]$consensusClass,results_pam[[10]]$consensusClass)
colnames(varGenes_combined.classes) = c("hierarchy", "partition")


# Using random forest, only on variable genes
x = resid[which(rownames(resid) %in% highvar2),]
varGenes.rf = randomForest(x=t(x),y=as.factor(varGenes_combined.classes[,2]),importance=TRUE, proximity=TRUE, ntree=1000)
imp = round(importance(varGenes.rf), 2)
barplot(abs(imp[order(abs(imp[,3]),decreasing=TRUE),3]), las = 2, main = "MeanDecAcc",names.arg = "")
barplot(imp[order(abs(imp[,3]),decreasing=TRUE),3], las = 2, main = "MeanDecAcc",names.arg = "")
barplot(imp[order(imp[,4],decreasing=TRUE),4], las = 2, main = "IncNodePurity",names.arg = "") 
varImpPlot(varGenes.rf)
varImpPlot(varGenes.rf,n.var = 150)
print(varGenes.rf)
varGenes.centers = classCenter(x=t(x),label=as.factor(varGenes_combined.classes[,2]),prox=varGenes.rf$proximity)
heatmap.2(varGenes.centers,labCol="",trace='n',col=bluered,scale='r')

head(imp)
length(which(imp[,11] > 3))
write.table(rownames(imp),file = "vargenes_list_10grp_gt3.txt",quote = FALSE,sep = "\n",row.names = FALSE,col.names = FALSE)


# How are the 2 groups distributed in the 10 groups?
twoGroupSplit = cbind(results_pam[[2]]$consensusClass, results_pam[[10]]$consensusClass)
pie(table(twoGroupSplit[,2]))
for (i in 1:10){
  print(paste("group", i, sep = " "))
  print(table(twoGroupSplit[which(twoGroupSplit[,2] == i),1]))
}
```
