---
title: "correlation_analysis.Rmd"
author: "Kristen Dang"
date: "7/18/2017"
output: html_document
---

```{r message=FALSE,warning=FALSE,echo=FALSE}
library(synapseClient)
synapseLogin()
library(gplots)
library(MASS)
library("VennDiagram")
library('githubr')
library('RColorBrewer')
library(ConsensusClusterPlus)


source('/Users/kkdang/Computing/rgithubclient_authenticate.R')
sourceRepoFile(sageCode, 'biomart_fxns.R')
setwd('~/Computing/cranio/IGF1-RUNX2-TWIST1/')
```


# Get data
```{r}
resid = read.delim(getFileLocation(synGet("syn8555302")),row.names = 1)
resid[1:5,1:6]
```


# Genes of Interest
```{r}
genesOfInterest = c("RUNX2", "IGF1", "TWIST1", "GSK3b")
geneNames = getBM(attributes=c("ensembl_gene_id","hgnc_symbol"),filters="hgnc_symbol",values=genesOfInterest,mart=Hs)    
print(geneNames)

temp = data.matrix(resid[which(rownames(resid) %in% geneNames$ensembl_gene_id),])
rownames(temp) = geneNames$hgnc_symbol[match(rownames(temp),geneNames$ensembl_gene_id)]


heatmap.2(temp,trace="none",scale="none", col=bluered,cexRow = 0.9)
```


# Overall data dist and filtering
```{r}
heatmap.2(data.matrix(resid[1:100,]),trace = "none",scale = "none",col=bluered)
allSD = apply(resid,MARGIN = 1,FUN = sd)
hist(allSD,breaks = 50)
truehist(allSD)
sum(allSD > 0.5)
highVar = which(allSD > 0.5)


intersect(geneNames$ensembl_gene_id, names(highVar))

allSD[which(names(allSD) == geneNames$ensembl_gene_id[1])]
```



# What is the signal in RUNX2, TWIST1, GSK3B for patients with high/low IGF1?
```{r}
results_pam = ConsensusClusterPlus(d=temp,maxK = 6,reps = 50,clusterAlg = "pam",distance = "pearson",pFeature = 1,writeTable=TRUE,plot="pdf",title="consCluster_4genes_pam_pear_pF1")
save(results_pam,file = "consClus_results_4gene_pam_p1.Robj.bz2",compress = "bzip2")


# Plot median gene signature for each group
gp1 = names(which(results_pam[[3]]$consensusClass == 1))
gp1Signature = apply(temp[,colnames(temp) %in% gp1],MARGIN = 1,median)
gp2 = names(which(results_pam[[3]]$consensusClass == 2))
gp2Signature = apply(temp[,colnames(temp) %in% gp2],MARGIN = 1,median)
gp3 = names(which(results_pam[[3]]$consensusClass == 3))
gp3Signature = apply(temp[,colnames(temp) %in% gp3],MARGIN = 1,median)
dotchart(t(cbind(gp1Signature,gp2Signature,gp3Signature)),xlab = "residualized expression")

```


# Getting diagnosis and other metadata
```{r}
## Filter and convert metadata
sourceRepoFile(sageCode, "scri_cran/process_metadata.R")
metadataFiltered = processMetadata(ver = 5,plot = FALSE)
metadataFiltered$Sample_Type = as.character(metadataFiltered$Sample_Type)
metadataFiltered$Sample_Type[grep("Coronal",metadataFiltered$Sample_Type)] = "Coronal"
metadataFiltered$Sample_Type[grep("Metopic",metadataFiltered$Sample_Type)] = "Metopic"
metadataFiltered$Sample_Type[grep("Sagittal",metadataFiltered$Sample_Type)] = "Sagittal"
metadataFiltered$Sample_Type = as.factor(metadataFiltered$Sample_Type)

editedColNames = substr(colnames(resid),start = 2,stop = 8)
ncol(resid)
colDiag = metadataFiltered$Sample_Type[match(editedColNames,metadataFiltered$SAMPLE)]
length(colDiag)
levels(colDiag)
```


# Correlation: what is the expression profile of correlates of RUNX2, TWIST1, IGF1?
```{r fig.height=8,fig.width=8}
allCors = cor(t(resid[highVar,]))
hist(allCors[lower.tri(allCors)])
truehist(allCors[lower.tri(allCors)])
diagColors = brewer.pal(n = 5, name = "Accent")

testCorrelationWith=function(inGene,adjMethod="BH") {
  results = apply(data.matrix(resid[highVar,]),MARGIN = 1,function(x){ cor.test(x,inGene,method = "pearson")} )
  allP = p.adjust(p = sapply(results,function(x) x$p.value),method = adjMethod)
  allEst = sapply(results,function(x) x$estimate)
  return(list(pvalue=allP,correlations=allEst))
}  

IGF1correlates = testCorrelationWith(inGene = t(resid[rownames(resid) == geneNames$ensembl_gene_id[2],]))
hist(IGF1correlates$pvalue)
truehist(IGF1correlates$pvalue)
truehist(IGF1correlates$correlations)
IGFgroup = names(IGF1correlates$pvalue)[intersect(which(IGF1correlates$pvalue < 0.05), which(abs(IGF1correlates$correlations) > 0.5))]


labRows = rownames(resid)[rownames(resid) %in% IGFgroup]
tokeep = which(labRows == geneNames$ensembl_gene_id[geneNames$hgnc_symbol == "IGF1"])
dispLab = rep("", length(IGFgroup))
dispLab[tokeep] = "IGF1"
heatmap.2(data.matrix(resid[rownames(resid) %in% IGFgroup,]),trace = "none",scale = "none",col = bluered,labRow = dispLab,labCol = "",main = "IGF1 correlates",ColSideColors = diagColors[as.numeric(colDiag)],key = TRUE )
legend("bottomleft",inset = .01,legend = levels(colDiag),pch = 15, col = diagColors)


# TWIST
TWIST1correlates = testCorrelationWith(inGene = t(resid[rownames(resid) == geneNames$ensembl_gene_id[4],]))
hist(TWIST1correlates$pvalue)
truehist(TWIST1correlates$pvalue)
truehist(TWIST1correlates$correlations)
TWISTgroup = names(TWIST1correlates$pvalue)[intersect(which(TWIST1correlates$pvalue < 0.05), which(abs(TWIST1correlates$correlations) > 0.5))]

labRows = rownames(resid)[rownames(resid) %in% TWISTgroup]
tokeep = which(labRows == geneNames$ensembl_gene_id[geneNames$hgnc_symbol == "TWIST1"])
dispLab = rep("", length(TWISTgroup))
dispLab[tokeep] = "TWIST1"
heatmap.2(data.matrix(resid[rownames(resid) %in% TWISTgroup,]),trace = "none",scale = "none",col = bluered,labRow = dispLab,labCol = "",main = "TWIST1 correlates",ColSideColors = diagColors[as.numeric(colDiag)],key = TRUE )
legend("bottomleft",inset = .01,legend = levels(colDiag),pch = 15, col = diagColors)


#RUNX2
RUNX2correlates = testCorrelationWith(inGene = t(resid[rownames(resid) == geneNames$ensembl_gene_id[3],]))
hist(RUNX2correlates$pvalue)
truehist(RUNX2correlates$pvalue)
truehist(RUNX2correlates$correlations)
RUNXgroup = names(RUNX2correlates$pvalue)[intersect(which(RUNX2correlates$pvalue < 0.05), which(abs(RUNX2correlates$correlations) > 0.5))]

labRows = rownames(resid)[rownames(resid) %in% RUNXgroup]
tokeep = which(labRows == geneNames$ensembl_gene_id[geneNames$hgnc_symbol == "RUNX2"])
dispLab = rep("", length(RUNXgroup))
dispLab[tokeep] = "RUNX2"
heatmap.2(data.matrix(resid[rownames(resid) %in% RUNXgroup,]),trace = "none",scale = "none",col = bluered,labRow = dispLab,labCol = "",main = "RUNX2 correlates",ColSideColors = diagColors[as.numeric(colDiag)],key = TRUE )
legend("bottomleft",inset = .01,legend = levels(colDiag),pch = 15, col = diagColors)

```



# List the genes that correlate with each 
```{r}
IGFgroupHGNC = getHGNC(inENSG = IGFgroup)
print(unique(IGFgroupHGNC$hgnc_symbol))


TWISTgroupHGNC = getHGNC(inENSG = TWISTgroup)
print(unique(TWISTgroupHGNC$hgnc_symbol))

RUNXgroupHGNC = getHGNC(inENSG = RUNXgroup)
print(unique(RUNXgroupHGNC$hgnc_symbol))

venn.diagram(x = list("IGF1"=IGFgroupHGNC$hgnc_symbol,"TWIST"=TWISTgroupHGNC$hgnc_symbol,"RUNX2"=RUNXgroupHGNC$hgnc_symbol),filename = "venn-3gene-analysis.tiff",fill = c("cornflowerblue", "green", "yellow"))

print(intersect(intersect(IGFgroupHGNC$hgnc_symbol,TWISTgroupHGNC$hgnc_symbol),RUNXgroupHGNC$hgnc_symbol))
```


# Do patients with strong correlate profiles in IGF1 overlap with patients with strong correlate signals in [RUNX2|TWIST1]?
```{r}
gpColors = c("tomato", "dodgerblue", "gold", "gray")

results_pam = ConsensusClusterPlus(d=data.matrix(resid[which(rownames(resid) %in% IGFgroup),]),pItem=0.9,maxK = 8,reps = 100,clusterAlg = "pam",distance = "pearson",pFeature = 0.7,writeTable=TRUE,plot="pdf",title="consCluster_IGFgenes_pam_pF7_pI9")
save(results_pam,file = "consCluster_IGFgenes_pam_pF7_pI9.Robj.bz2",compress = "bzip2")
IGF3gp = results_pam[[3]]$consensusClass

heatmap.2(data.matrix(resid[rownames(resid) %in% IGFgroup,]),trace = "none",scale = "none",col = bluered,labRow = "",labCol = "",main = "IGF1 correlates",ColSideColors = gpColors[IGF3gp],key = TRUE )
legend("bottomleft",inset = .01,legend = c("1", "2", "3"),pch = 15, col = gpColors)



# TWIST1
results_pam = ConsensusClusterPlus(d=data.matrix(resid[which(rownames(resid) %in% TWISTgroup),]),pItem=0.9,maxK = 6,reps = 100,clusterAlg = "pam",distance = "pearson",pFeature = 0.7,writeTable=TRUE,plot="pdf",title="consCluster_TWISTgenes_pam_pF7_pI9")
save(results_pam,file = "consCluster_TWISTgenes_pam_pF7_pI9.Robj.bz2",compress = "bzip2")
TWIST4gp = results_pam[[4]]$consensusClass

heatmap.2(data.matrix(resid[rownames(resid) %in% TWISTgroup,]),trace = "none",scale = "none",col = bluered,labRow = "",labCol = "",main = "TWIST1 correlates",ColSideColors = gpColors[TWIST4gp],key = TRUE )
legend("bottomleft",inset = .01,legend = c("1", "2", "3","4"),pch = 15, col = gpColors)

# RUNX1
results_pam = ConsensusClusterPlus(d=data.matrix(resid[which(rownames(resid) %in% RUNXgroup),]),pItem=0.9,maxK = 6,reps = 100,clusterAlg = "pam",distance = "pearson",pFeature = 0.7,writeTable=TRUE,plot="pdf",title="consCluster_RUNXgenes_pam_pF7_pI9")
save(results_pam,file = "consCluster_RUNXgenes_pam_pF7_pI9.Robj.bz2",compress = "bzip2")
RUNX2gp = results_pam[[2]]$consensusClass

heatmap.2(data.matrix(resid[rownames(resid) %in% RUNXgroup,]),trace = "none",scale = "none",col = bluered,labRow = "",labCol = "",main = "RUNX1 correlates",ColSideColors = gpColors[RUNX2gp],key = TRUE )
legend("bottomleft",inset = .01,legend = c("1", "2"),pch = 15, col = gpColors)


venn.diagram(list(RUNX2=names(RUNX2gp[RUNX2gp == 1]),TWIST1=names(TWIST4gp[TWIST4gp==1]),IGF1=names(IGF3gp[IGF3gp==1 | IGF3gp == 2])),filename = "venn-samps-strongvar.tiff",fill = c("cornflowerblue", "green", "yellow"))
```

# Classify patients into groups based on signal of correlates of IGF1, RUNX2, TWIST1?
### Would need to iteratively choose the cutoff for correlated genes to include.
### Show whether there is a distinct diagnostic group based on these genes, separate from controls.


# Plot / ANOVA of differences in mean expression per diagnostic group 

