---
title: "swap_correlation_and_DE.Rmd"
author: "Kristen Dang"
date: "Jan. 18, 2016"
output: 
  html_document: 
    fig_height: 7
---



```{r}
library('githubr')
source('/Users/kkdang/Computing/rgithubclient_authenticate.R')
library(edgeR)
library(gplots)
library(RColorBrewer)
library(nlme)
library(lme4)
sourceRepoFile(sageCode, "rnaseq_analysis_functions.R")
```

# Get data
```{r cache=TRUE}
sourceRepoFile(sageCode, "NIA-AMP-AD/swap_analysis/mssm_swap_preliminary.R")
# Fix mayo file names so they match the metadata
match(colnames(mssmR.dge), metadata$Sample.ID)
colnames(mayoR.dge)
x = paste("BM_10_", substr(colnames(mayoR.dge),start = 4,6), sep = "")
colnames(mayoR.dge) = x
colnames(mayoR.dge)
x = paste("BM_10_", substr(colnames(mayoPalx.dge),start = 4,6), sep = "")
colnames(mayoPalx.dge) = x
colnames(mayoPalx.dge)
colnames(mssmR.dge)

x = paste("BM_10_", substr(colnames(mayoR_counts),start = 4,6), sep = "")
colnames(mayoR_counts) = x
colnames(mayoR_counts)

x = paste("BM_10_", substr(colnames(mayoCountsPalx),start = 4,6), sep = "")
colnames(mayoCountsPalx) = x
colnames(mayoCountsPalx)

metadata$SEX = as.factor(metadata$SEX)
```


# Correlation analysis of processed data:
```{r}
# Get common set of genes to use
mssm_mayo = intersect(rownames(mssmR_counts),rownames(mayoR_counts))
mssm_Common = mssmR_counts[which(rownames(mssmR_counts) %in% mssm_mayo),]
mssm_Common = mssm_Common[order(rownames(mssm_Common)),]
mayo_Common = mayoR_counts[which(rownames(mayoR_counts) %in% mssm_mayo),]
mayo_Common = mayo_Common[order(rownames(mayo_Common)),]

tail(rownames(mssm_Common))
tail(rownames(mayo_Common))
dim(mssm_Common)
dim(mayo_Common)



mssm_mayoP = intersect(rownames(mssmCountsPalx),rownames(mayoCountsPalx))
mssm_CommonP = mssmCountsPalx[which(rownames(mssmCountsPalx) %in% mssm_mayoP),]
mssm_CommonP = mssm_CommonP[order(rownames(mssm_CommonP)),]
mayo_CommonP = mayoCountsPalx[which(rownames(mayoCountsPalx) %in% mssm_mayoP),]
mayo_CommonP = mayo_CommonP[order(rownames(mayo_CommonP)),]

tail(rownames(mssm_CommonP))
tail(rownames(mayo_CommonP))
dim(mssm_CommonP)
dim(mayo_CommonP)
```

How well do samples correlate to each other across sites? Plot distribution of per-sample correlations across sites. Compare to correlation within sites.

1. Mayo-MSSM

```{r}
mssm_Common = mssm_Common[,match(colnames(mayo_Common),colnames(mssm_Common))]
colnames(mssm_Common)
colnames(mayo_Common)
mm_Corr = cor(mssm_Common,mayo_Common,method = "spearman")
heatmap.2(mm_Corr,trace="none",margins = c(7,7), main = "Mayo-MSSM",breaks = seq(0.8,0.95,0.01))

mssm_CommonP = mssm_CommonP[,match(colnames(mayo_CommonP),colnames(mssm_CommonP))]
colnames(mssm_CommonP)
colnames(mayo_CommonP)
mm_CorrP = cor(mssm_CommonP,mayo_CommonP,method = "spearman")
heatmap.2(mm_CorrP,trace="none",margins = c(7,7), main = "Mayo-MSSM PALX",breaks = seq(0.8,0.95,0.01))
```

# All together now
```{r}
plot(density(mm_Corr), ylim = range(0,45), main = "Mayo-MSSM")
lines(density(diag((mm_Corr))),lty = 2)
legend("topright",legend = c("all", "self"),lty = c(1,2))

plot(density(mm_CorrP), ylim = range(0,45), main = "Mayo-MSSM PALX")
lines(density(diag((mm_CorrP))),lty = 2)
legend("topright",legend = c("all", "self"),lty = c(1,2))

removeSelf=function(inSelf){
  selfMatches = diag(inSelf)
  temp = cbind(diag(x=selfMatches),matrix(0,nrow=9,ncol = 0))
  return(inSelf - temp)
}
boxplot(as.vector(removeSelf(mm_Corr)), diag(mm_Corr), varwidth = TRUE,ylim = range(0.75,1), col = c("goldenrod1","goldenrod1"),names = c("non-self","self"), main = "all genes")
boxplot(as.vector(removeSelf(mm_CorrP)), diag(mm_CorrP), varwidth = TRUE,ylim = range(0.75,1), col = c("goldenrod1","goldenrod1"),names = c("non-self","self"), main = "filtered")
```

PCA, sample clustering, etc.
```{r}
mssm_temp.dge = mssmPalx.dge[which(rownames(mssmPalx.dge) %in% mssm_mayoP),]
mssm_temp.dge = mssm_temp.dge[order(rownames(mssm_temp.dge)),]
colnames(mssm_temp.dge) = paste("M",colnames(mssm_temp.dge),sep = "_")
mssm_temp.dge = mssm_temp.dge[,order(colnames(mssm_temp.dge))]
mayo_temp.dge = mayoPalx.dge[which(rownames(mayoPalx.dge) %in% mssm_mayoP),]
mayo_temp.dge = mayo_temp.dge[order(rownames(mayo_temp.dge)),]
colnames(mayo_temp.dge) = paste("Y",colnames(mayo_temp.dge),sep = "_")
mayo_temp.dge = mayo_temp.dge[,order(colnames(mayo_temp.dge))]
all.dge = DGEList(counts = cbind(getCounts(mssm_temp.dge),getCounts(mayo_temp.dge)))
all.dge = calcNormFactors(all.dge)

fm.pca = FM_PCA(in.data = cpm(all.dge,normalized.lib.sizes=TRUE,log=TRUE),plot = TRUE)
#write.table(fm.pca$ind$coord,file = "PCs_atMSSM.tsv",quote = FALSE,sep = "\t",row.names = TRUE,col.names = TRUE)


Dx = as.numeric(as.factor(rep(metadata$CDR[match(substr(colnames(mssm_temp.dge),start = 3,stop = 15),metadata$Sample.ID)],2)))
Dx = Dx[-16]
DxColors = brewer.pal(5, 'YlOrRd')
SampColors = rep(brewer.pal(n=10,name="Paired"),2)
SampColors = SampColors[-16]
plot.PCA(fm.pca,choix = "ind",new.plot = TRUE,label="all",cex=0.8)

op = par(bg = "gray")
plot.PCA(fm.pca,choix = "ind",new.plot = TRUE,label="all",cex=0.8,col.ind = DxColors[Dx], axes = c(2,3))
legend("bottomright",legend = c(0,1,2,3,5),pch = 16, col = DxColors )
par(op)
plot.PCA(fm.pca,choix = "ind",new.plot = TRUE,label="all",cex=0.8,axes = c(2,3),col.ind = SampColors)


op = par(bg = "gray")
plot.PCA(fm.pca,choix = "ind",new.plot = TRUE,label="all",cex=0.8,col.ind = DxColors[Dx], axes = c(3,4))
legend("bottomright",legend = c(0,1,2,3,5),pch = 16, col = DxColors )
par(op)
plot.PCA(fm.pca,choix = "ind",new.plot = TRUE,label="all",cex=0.8,axes = c(3,4),col.ind = SampColors)


op = par(bg = "gray")
plot.PCA(fm.pca,choix = "ind",new.plot = TRUE,label="all",cex=0.8,axes = c(4,5),col.ind = DxColors[Dx])
legend("bottomright",legend = c(0,1,2,3,5),pch = 16, col = DxColors )
par(op)
plot.PCA(fm.pca,choix = "ind",new.plot = TRUE,label="all",cex=0.8,axes = c(4,5),col.ind = SampColors)

```


Model selection comparing a series of non-nested, 1-factor models. 
```{r}
metadata$CDR = as.factor(metadata$CDR)

combinedMeta = rbind(metadata[match(colnames(mssmR.dge), metadata$Sample.ID),], metadata[match(colnames(mssmR.dge), metadata$Sample.ID),])
combinedMeta$site = c(rep("MSSM",10),rep("Mayo",10))
combinedMeta$CERJ = as.factor(combinedMeta$CERJ)
combinedMeta$CDR = as.factor(combinedMeta$CDR)
combinedMeta$SEX = as.factor(combinedMeta$SEX)
combinedMeta$Donor = as.factor(combinedMeta$Donor)
combinedMeta$NP.1 = as.factor(combinedMeta$NP.1)
combinedMeta$bbscore = as.factor(combinedMeta$bbscore)
combinedMeta$site = as.factor(combinedMeta$site)
combinedMeta$libsize = log10(all.dge$samples$lib.size[match(paste(c(rep("M_", 10), rep("Y_", 10)), combinedMeta$Sample.ID, sep = ""), rownames(all.dge$samples))])

toRemove = setdiff(combinedMeta$Sample.ID, colnames(mayoR.dge))
x = which(combinedMeta$Sample.ID == toRemove)
combinedMeta = combinedMeta[-x[2],]

design = model.matrix(~CDR+CERJ+SEX+PMI+NP.1+bbscore+site+libsize,data = combinedMeta)
rownames(design) = c(paste("M_",combinedMeta$Sample.ID[1:10],sep = ""),paste("Y_",combinedMeta$Sample.ID[11:19],sep = ""))
all.dge = all.dge[,match(colnames(all.dge),rownames(design))]

makeModels=function(){

  CDR = model.matrix(~CDR,data = combinedMeta)
  CERJ = model.matrix(~CERJ,data = combinedMeta)
  SEX = model.matrix(~SEX,data = combinedMeta)
#  Age_death = model.matrix(~Age_death,data = combinedMeta)
  bbscore = model.matrix(~bbscore,data = combinedMeta) 
  PMI = model.matrix(~PMI,data = combinedMeta)
  NP = model.matrix(~NP.1,data = combinedMeta)
  site = model.matrix(~site,data = combinedMeta)
  libsize = model.matrix(~libsize, data = combinedMeta)

  designlist = list(CDR=CDR,CERJ=CERJ,SEX=SEX,bbscore=bbscore,NP=NP,PMI=PMI,site=site,libsize=libsize)
  return(designlist)
}

allModels=makeModels()
chooseModel = selectModel(voom(all.dge),allModels,criterion = "bic")
x = table(chooseModel$pref)
chooseModel = selectModel(voom(all.dge),allModels,criterion = "aic")
votes = rbind(x,table(chooseModel$pref))
barplot(votes,beside=TRUE, ylab = "# of genes choosing this model",legend.text = c("BIC", "AIC"),main="model selection for transcripts", cex.names = 0.7, las = 2)
barplot(votes,beside=TRUE, ylab = "# of genes choosing this model",legend.text = c("BIC", "AIC"),main="model selection for transcripts", cex.names = 0.7, las = 2,ylim = range(0,5000))
```

ANOVA of count data. For each transcript, the percentage of variation explained by each factor is calculated dividing the sum of squares by the total sum of squares. 
```{r echo=FALSE,eval=FALSE}
# Using a linear model without accounting for repeated measures
fracVar  = matrix(NA,nrow = nrow(all.dge),ncol=length(allModels))
all.cpm = cpm(all.dge,normalized.lib.sizes = TRUE,log = TRUE)
for (i in 1:nrow(all.cpm)){
  geneFit = lm(all.cpm[i,]~CDR+CERJ+SEX+bbscore+PMI+site+libsize,data = combinedMeta)
  x = summary(aov(geneFit))
  fracVar[i,] = x[[1]]$`Sum Sq` / sum(x[[1]]$`Sum Sq`)
}
op = par(mar=c(7,4,4,2)+0.1)
boxplot(fracVar,names=rownames(x[[1]]),las = 2)
boxplot(fracVar,names=rownames(x[[1]]),las = 2,ylim = range(0,0.3),outline =FALSE)
par(op)


# lme4 
# Using a mixed model with subject as random (block) and site as fixed
fracVar  = matrix(NA,nrow = nrow(all.dge),ncol=length(allModels))
all.cpm = cpm(all.dge,normalized.lib.sizes = TRUE,log = TRUE)
for (i in 1:nrow(all.cpm)){
  geneFit = lmer(all.cpm[i,]~CDR+SEX+site+log(libsize) + (1|Sample.ID),data = combinedMeta)
  x = anova(geneFit)
  fracVar[i,] = x$`Sum Sq` / sum(x$`Sum Sq`)
}
boxplot(fracVar[,1:4],names=rownames(x),las = 2, ylim = range(0,1), ylab = "sum sq / sum(sum sq)")
par(op)





# Testing linear-mixed model with repeated measures in one gene
geneFit.gd = groupedData(all.cpm[i,] ~ site + libsize | Sample.ID, data=cbind(combinedMeta, all.cpm[i,]), outer = ~ CDR+CERJ+SEX+bbscore+PMI, labels = list(y="expression"))
x = which(colnames(geneFit.gd) %in% "all.cpm[i, ]")
colnames(geneFit.gd)[x] = "expression"
plot(geneFit.gd)
geneFit2 = lme(fixed = expression ~ site+CDR+SEX+PMI+PlaqueMean+libsize, data = geneFit.gd, random = ~ libsize | Sample.ID)
geneFit3 = update(geneFit2, random = ~ 1 | Sample.ID)
anova(geneFit2, geneFit3)
geneFit4 = update(geneFit2, random = ~ PMI | Sample.ID)
anova(geneFit4, geneFit3)
geneFit5 = update(geneFit2, random = ~ PlaqueMean | Sample.ID)
anova(geneFit5, geneFit3)


# Using lme with repeated measures
elements = c("site","CDR","SEX","PMI","PlaqueMean","libsize")
cAIC = matrix(NA, nrow = nrow(all.dge), ncol=(length(elements)+1))
cBIC = matrix(NA, nrow = nrow(all.dge), ncol=(length(elements)+1))
all.cpm = cpm(all.dge,normalized.lib.sizes = TRUE,log = TRUE)
for (i in 1:nrow(all.cpm)){
  geneFit.gd = groupedData(all.cpm[i,] ~ site + libsize | Sample.ID, data=cbind(combinedMeta, all.cpm[i,]), outer = ~ CDR+CERJ+SEX+bbscore+PMI, labels = list(y="expression"))
   x = which(colnames(geneFit.gd) %in% "all.cpm[i, ]")
  colnames(geneFit.gd)[x] = "expression"
  try({
    geneFit.full = lme(fixed = as.formula(paste("expression ~",paste(elements, collapse = "+"),sep = " ")), data = geneFit.gd, random = ~ 1 | Sample.ID)
    for (j in 1:length(elements)) {
      testFormula = as.formula(paste("expression ~",paste(elements[-j], collapse = "+"),sep = " "))
      try({
        geneFit.up = update(geneFit.full, fixed = testFormula)
        cAIC[i,j] = summary(geneFit.up)$AIC
        cBIC[i,j] = summary(geneFit.up)$BIC
      },silent = T)
    }
    cAIC[i,(length(elements)+1)] = summary(geneFit.full)$AIC
    cBIC[i,(length(elements)+1)] = summary(geneFit.full)$BIC
  },silent = T)
}
op = par(mar=c(7,4,4,2)+0.1)
boxplot(cAIC, names = c(elements, "full"), las = 2, main = "model comparison with AIC")
boxplot(cBIC, names = c(elements, "full"), las = 2, main = "model comparison with BIC")
#boxplot(fracVar,names=rownames(x[[1]]),las = 2,ylim = range(0,0.3),outline =FALSE)
par(op)

```
